resistanceModelNotesAndy

### 2016-03-11 paper1_results_figs.Rmd
tweak figs

  df_diff <- cbind(resistPointsMix_A_T[-indices1], tmp)

  #remove runs that didn't reach a resistance threshold (999), even if just for one strategy
  #>1000 excludes sequential strategy that had a 999 in
  #BEWARE would need to increase 1000 if I increase max generations in the runs
  didntReachThresh <- which( df_diff$gen_cP0.25 > 500 | df_diff$gen_cP0.25seq > 500 )

  #subset by runs not making threshold
  df_diff <- df_diff[-didntReachThresh,]

  #df_diff$mixA_minus_seq0.25 <- df_diff["gen_cP0.25"] - df_diff["gen_cP0.25seq"]
  df_diff["mixA_minus_seq0.25"] <- df_diff["gen_cP0.25"] - df_diff["gen_cP0.25seq"]

  #TODO : later do PRCC on df_diff["mixA_minus_seq0.25"]

@@

  ggSubset <- df_diff

@@

  ggSubset <- df_diff
  #if mix-seq >1 resistance took longer for mix so mix is better
  ggSubset$mixA_better_seq0.25 <- ifelse( ggSubset$mixA_minus_seq0.25 > 0, 1, 0 )
  ggSubset$seq_better_mixA0.25 <- ifelse( ggSubset$mixA_minus_seq0.25 < 0, 1, 0 )
  ggSubset$abs_diff <- abs(ggSubset$mixA_minus_seq0.25)

  #tricky
  #be a bit careful what to do with zeros where results are same for both strategies
  #I use these better vars to facet the plots
  #may want the zeroes in both facets
  #to get that I can rbind the zeros from one onto the other BEWARE this may screw up existing plots
  df_same <- ggSubset[which(ggSubset$abs_diff==0),]
  #set these to 1 (they were 0)
  df_same$mixA_better_seq0.25 <- 1
  df_same$seq_better_mixA0.25 <- 1
  #rbind them back onto other dataset
  ggSubset <- rbind( ggSubset, df_same)

  #takes a long time to plot, initial version gave me count
  #I want it to plot the mean
  #check stackoverflow ..

    print( ggplot(ggSubset, aes_string(x='exposure', y='effectiveness_insecticide1', colour='mixA_better_seq0.25'))+
             geom_hex()
             #points not wanted if 10000
             #geom_point(shape=1, colour="darkblue", alpha=0.2, show.legend=FALSE) + 
             #geom_smooth(colour='red', linetype='dashed',size=0.5) +
             #geom_smooth(linetype='dashed',size=1, colour="red") +
             #facet_wrap( ~ strategy) +
             #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
             #labs(title = i) +
             #geom_hline(yintercept = 0, colour="white")
           )

@@

    #COOLIO !! I really like this figure.
    library(viridis)
    #facet showing generation differences, labelled mix better, seq better
    #+ve & -ve on the same colour scale
    #failed attempt at getting Curtis point in legend
    #print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_diff, size="Curtis"))+
    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_diff))+

@@

    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_diff))+

@@

             scale_color_viridis(begin=1, end=0, option="viridis", 
                                 guide=guide_colourbar(title="difference in \ngenerations to reach\nresistance threshold")) + 
             theme_bw() +
             #add point for Curtis inputs
             geom_point(data = curtisInputs, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2), colour='red', shape=3) +
             ylab('effectiveness of insecticides added') +
             xlab('exposure to both insecticides') + 
             #main points for the plot
             geom_point(alpha=1, size=0.5) + 
             #!!!NOW can I add a lm fitted to the 0 points abs_diff
             geom_smooth(data=ggSubset[which(ggSubset$abs_diff==0),], method = "lm", se = FALSE) +
             #can I subtley highlight the zero points which are repeated in each plot ?
             geom_point(data=ggSubset[which(ggSubset$abs_diff==0),], size=0.8, colour='lightgrey', shape=2, fill='white') +            
             #add text for Curtis point
             geom_text(data = curtisInputs, inherit.aes = FALSE, aes_string(x = 'exposure', y='effectiveness_insecticide1 + effectiveness_insecticide2', label='"Curtis Fig.2"'), nudge_x=-0.05, size=3, colour='red')
           )

# doing the lm outside of ggplot    
df_same = ggSubset[which(ggSubset$abs_diff==0),]
x1=df_same$exposure 
y1=df_same$effectiveness_insecticide1 + df_same$effectiveness_insecticide2

@@

x2 <- seq(0,1,0.1)
#approximation
y2 <- 0.6*x2+0.98
plot(x2,y2)

@@

    x <- df_diff[,1:num_inputs]
    y <- df_diff["mixA_minus_seq0.25"]

>>>>

  dif_mixA_seq <- cbind(resistPointsMix_A_T[-indices1], tmp)
  dif_mix2_seq <- cbind(resistPointsMix_2_T[-indices1], tmp)

  #remove runs that didn't reach a resistance threshold (999), even if just for one strategy
  #>1000 excludes sequential strategy that had a 999 in
  #BEWARE would need to increase 1000 if I increase max generations in the runs
  didntReachThresh <- which( dif_mixA_seq$gen_cP0.25 > 500 | dif_mixA_seq$gen_cP0.25seq > 500 )
  #subset by runs not making threshold
  dif_mixA_seq <- dif_mixA_seq[-didntReachThresh,]

  dif_mixA_seq["mixA_minus_seq0.25"] <- dif_mixA_seq["gen_cP0.25"] - dif_mixA_seq["gen_cP0.25seq"]

  #for dif_mix2_seq
  didntReachThresh <- which( dif_mix2_seq$gen_cP0.25 > 500 | dif_mix2_seq$gen_cP0.25seq > 500 )
  #subset by runs not making threshold
  dif_mix2_seq <- dif_mix2_seq[-didntReachThresh,]

  dif_mix2_seq["mix2_minus_seq0.25"] <- dif_mix2_seq["gen_cP0.25"] - dif_mix2_seq["gen_cP0.25seq"]

@@

  ggSubset <- dif_mixA_seq

@@

  ##BEWARE toogle this to look at plots for mixboth vs seq
  ggSubset <- dif_mixA_seq
  #ggSubset <- dif_mix2_seq

  #if mix-seq >1 resistance took longer for mix so mix is better
  ggSubset$mixA_better_seq0.25 <- ifelse( ggSubset$mixA_minus_seq0.25 > 0, 1, 0 )
  ggSubset$seq_better_mixA0.25 <- ifelse( ggSubset$mixA_minus_seq0.25 < 0, 1, 0 )
  ggSubset$abs_mixA_minus_seq0.25 <- abs(ggSubset$mixA_minus_seq0.25)

  #tricky
  #be a bit careful what to do with zeros where results are same for both strategies
  #I use these better vars to facet the plots
  #may want the zeroes in both facets
  #to get that I can rbind the zeros from one onto the other BEWARE this may screw up existing plots
  df_same <- ggSubset[which(ggSubset$mixA_minus_seq0.25==0),]
  #set these to 1 (they were 0)
  df_same$mixA_better_seq0.25 <- 1
  df_same$seq_better_mixA0.25 <- 1
  #rbind them back onto other dataset, probably not needed ??
  ggSubset <- rbind( ggSubset, df_same)

  #add an option for 20% better, put the not in to get facets in correct order in ggplot later
  ggSubset$mixA_not20pc_better_seq0.25 <- ifelse( ggSubset$mixA_minus_seq0.25 / ggSubset$gen_cP0.25seq <= 0.2, 1, 0 )

    #failed hex plot just gave me counts
    # print( ggplot(ggSubset, aes_string(x='exposure', y='effectiveness_insecticide1', colour='mixA_better_seq0.25'))+
    #          geom_hex()
    #        )

@@

    #COOLIO !! I really like this figure.
    library(viridis)
    #facet showing generation differences, labelled mix better, seq better
    #+ve & -ve on the same colour scale
    #failed attempt at getting Curtis point in legend
    #print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mixA_minus_seq0.25, size="Curtis"))+
    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mixA_minus_seq0.25))+

@@

    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mixA_minus_seq0.25))+

@@

             scale_color_viridis(begin=1, end=0, option="viridis", 
                                 guide=guide_colourbar(title="difference in \ngenerations to reach\nresistance threshold")) + 
             theme_bw() +
             #add point for Curtis inputs
             geom_point(data = curtisInputs, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2), colour='red', shape=3) +
             ylab('effectiveness of insecticides added') +
             xlab('exposure to both insecticides') + 
             #main points for the plot
             geom_point(alpha=1, size=0.5) + 
             #!!!NOW can I add a lm fitted to the 0 points mixA_minus_seq0.25
             geom_smooth(data=ggSubset[which(ggSubset$mixA_minus_seq0.25==0),], method = "lm", se = FALSE) +
             #can I subtley highlight the zero points which are repeated in each plot ?
             geom_point(data=ggSubset[which(ggSubset$mixA_minus_seq0.25==0),], size=0.8, colour='lightgrey', shape=2, fill='white') +            
             #add text for Curtis point
             geom_text(data = curtisInputs, inherit.aes = FALSE, aes_string(x = 'exposure', y='effectiveness_insecticide1 + effectiveness_insecticide2', label='"Curtis Fig.2"'), nudge_x=-0.05, size=3, colour='red')
           )

  # doing the lm outside of ggplot    
  df_same = ggSubset[which(ggSubset$mixA_minus_seq0.25==0),]
  x1=df_same$exposure 
  y1=df_same$effectiveness_insecticide1 + df_same$effectiveness_insecticide2

@@

  x2 <- seq(0,1,0.1)
  #approximation
  y2 <- 0.6*x2+0.98
  #plot(x2,y2)

  #modify the killer plot to show where mixtures are 20% better ?
  print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2))+ #, colour=abs_mixA_minus_seq0.25))+

             facet_grid( ~ mixA_not20pc_better_seq0.25, labeller=as_labeller( c('0'='adaptive mix\nresistance >20% slower','1'='adaptive mix\nresistance <20% slower'))) +

             theme_bw() +
             #add point for Curtis inputs
             #geom_point(data = curtisInputs, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2), colour='red', shape=3) +
             ylab('effectiveness of insecticides added') +
             xlab('exposure to both insecticides') + 
             #main points for the plot
             geom_point(alpha=1, size=0.5) +
             #add a lm fitted to the 0 points
             geom_smooth(data=df_same, method = "lm", se = FALSE) 
           )

# repeating killer plot for mixtures both versus sequential
```{r, eval=TRUE, echo=FALSE, fig.height=3.5}

  #try repeating the killer plot for mixtures both versus sequential

  ##BEWARE toogle this to look at plots for mixboth vs seq
  #ggSubset <- dif_mixA_seq
  ggSubset <- dif_mix2_seq

  #if mix-seq >1 resistance took longer for mix so mix is better
  ggSubset$mix2_better_seq0.25 <- ifelse( ggSubset$mix2_minus_seq0.25 > 0, 1, 0 )
  ggSubset$seq_better_mix20.25 <- ifelse( ggSubset$mix2_minus_seq0.25 < 0, 1, 0 )
  ggSubset$abs_mix2_minus_seq0.25 <- abs(ggSubset$mix2_minus_seq0.25)

  #tricky
  #be a bit careful what to do with zeros where results are same for both strategies
  #I use these better vars to facet the plots
  #may want the zeroes in both facets
  #to get that I can rbind the zeros from one onto the other BEWARE this may screw up existing plots
  df_same <- ggSubset[which(ggSubset$mix2_minus_seq0.25==0),]
  #set these to 1 (they were 0)
  df_same$mix2_better_seq0.25 <- 1
  df_same$seq_better_mix20.25 <- 1
  #rbind them back onto other dataset, probably not needed ??
  ggSubset <- rbind( ggSubset, df_same)

  #add an option for 20% better, put the not in to get facets in correct order in ggplot later
  ggSubset$mix2_not20pc_better_seq0.25 <- ifelse( ggSubset$mix2_minus_seq0.25 / ggSubset$gen_cP0.25seq <= 0.2, 1, 0 )

    #i think adding the effectivenesses may give an even better plot
    #and makes it easier for the reader to interpret
    #aha! yes that's even better
    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=mix2_better_seq0.25))+

    #facetted 0/1, labelled mix better, seq better
    #if I get this sorted I could try to have +ve & -ve on the same scale
    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=seq_better_mix20.25))+
             facet_grid( ~ seq_better_mix20.25, labeller=as_labeller( c('0'='mix better','1'='sequential better'))) +

    #COOLIO !! I really like this figure.
    library(viridis)
    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mix2_minus_seq0.25))+

             facet_grid( ~ seq_better_mix20.25, labeller=as_labeller( c('0'='mix better','1'='sequential better'))) +

  #modify the killer plot to show where mixtures are 20% better ?
  print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2))+ #, colour=abs_mixA_minus_seq0.25))+

             facet_grid( ~ mix2_not20pc_better_seq0.25, labeller=as_labeller( c('0'='mix\nresistance >20% slower','1'='mix\nresistance <20% slower'))) +

             theme_bw() +
             #add point for Curtis inputs
             #geom_point(data = curtisInputs, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2), colour='red', shape=3) +
             ylab('effectiveness of insecticides added') +
             xlab('exposure to both insecticides') + 
             #main points for the plot
             geom_point(alpha=1, size=0.5) +
             #add a lm fitted to the 0 points
             geom_smooth(data=df_same, method = "lm", se = FALSE) 
           )

@@

    x <- dif_mixA_seq[,1:num_inputs]
    y <- dif_mixA_seq["mixA_minus_seq0.25"]

### 2016-03-11 paper1_results_figs_slimmed.Rmd
tweak figs

  df_diff <- cbind(resistPointsMix_A_T[-indices1], tmp)

  #remove runs that didn't reach a resistance threshold (999), even if just for one strategy
  #>1000 excludes sequential strategy that had a 999 in
  #BEWARE would need to increase 1000 if I increase max generations in the runs
  didntReachThresh <- which( df_diff$gen_cP0.25 > 500 | df_diff$gen_cP0.25seq > 500 )

  #subset by runs not making threshold
  df_diff <- df_diff[-didntReachThresh,]

  #df_diff$mixA_minus_seq0.25 <- df_diff["gen_cP0.25"] - df_diff["gen_cP0.25seq"]
  df_diff["mixA_minus_seq0.25"] <- df_diff["gen_cP0.25"] - df_diff["gen_cP0.25seq"]

  #TODO : later do PRCC on df_diff["mixA_minus_seq0.25"]

@@

  ggSubset <- df_diff
  #if mix-seq >1 resistance took longer for mix so mix is better
  ggSubset$mixA_better_seq0.25 <- ifelse( ggSubset$mixA_minus_seq0.25 > 0, 1, 0 )
  ggSubset$seq_better_mixA0.25 <- ifelse( ggSubset$mixA_minus_seq0.25 < 0, 1, 0 )
  ggSubset$abs_diff <- abs(ggSubset$mixA_minus_seq0.25)

>>>>

  dif_mixA_seq <- cbind(resistPointsMix_A_T[-indices1], tmp)

  #remove runs that didn't reach a resistance threshold (999), even if just for one strategy
  #>1000 excludes sequential strategy that had a 999 in
  #BEWARE would need to increase 1000 if I increase max generations in the runs
  didntReachThresh <- which( dif_mixA_seq$gen_cP0.25 > 500 | dif_mixA_seq$gen_cP0.25seq > 500 )

  #subset by runs not making threshold
  dif_mixA_seq <- dif_mixA_seq[-didntReachThresh,]

  #dif_mixA_seq$mixA_minus_seq0.25 <- dif_mixA_seq["gen_cP0.25"] - dif_mixA_seq["gen_cP0.25seq"]
  dif_mixA_seq["mixA_minus_seq0.25"] <- dif_mixA_seq["gen_cP0.25"] - dif_mixA_seq["gen_cP0.25seq"]

  #TODO : later do PRCC on dif_mixA_seq["mixA_minus_seq0.25"]

@@

  ggSubset <- dif_mixA_seq
  #if mix-seq >1 resistance took longer for mix so mix is better
  ggSubset$mixA_better_seq0.25 <- ifelse( ggSubset$mixA_minus_seq0.25 > 0, 1, 0 )
  ggSubset$seq_better_mixA0.25 <- ifelse( ggSubset$mixA_minus_seq0.25 < 0, 1, 0 )
  ggSubset$abs_diff <- abs(ggSubset$mixA_minus_seq0.25)

### 2016-03-11 resistanceModelNotesAndy.Rmd
tweak figs

msg from Ian :

Aha! I've understood what he means by setting correct_deploy to 0, to study mosaics.

~ changed order of strategies in legend

~ created new killer figure with added effectivenesses on the y & exposure on x.

** 8/3/2016 **

~ add Curtis fig2 params to killer figure
~ tricky adding to legend or labelling
~ can I get legend for killer figure beneath ?
~ can I derive a line for the border between when seq & mix better by fitting a line to the 0 values ?
~ get the classification tree that matches my killer figure

** 9/3/2016 **

~ creating 20% tree & plot
~ got the 0 line onto the first facet of the 20% plot

@@

@@

@@

~ Ian wants trees. Maybe just do one tree of resistance0.25 on whether seq or maxAdapt is better.

adaptive & sequential
20% longer

@@

~ I could try a 2D plot with x=exposure, y=effectiveness & hexbin & colour cells according to proportion of runs where mix or seq is better.

~~ seq is sometimes better than mix both. Where does that occur. The above plot might even help to answer that.

@@

## might be good to facet by mixture sequence better and just have points of x=exposure, y=effectiveness_insecticide1

@@

### My summary of results & thoughts

Resistance to two insecticides arises more slowly if they are used in combination rather than sequentially. See violin plots.

~ I'd like to show prcc on an axis from 0-1 with diff sysmbols for neg & pos. We care about which are most imp. params, not whether they are pos or neg.

~ If I would like to get prcc plots in a similar format to my existing plots. (but may not be space for axis labels)
~ get data in form : 
input1, input2, inputN, strategy 
prcc vals,,             indecticide1

@@

Main question is : Tree1 is mixture better (0/1), Tree2 is mixture better by >20%.

>>>>

msg from Ian :
Aha! I've understood what he means by setting correct_deploy to 0, to study mosaics.
~ changed order of strategies in legend
~ created new killer figure with added effectivenesses on the y & exposure on x.

** 8/3/2016 **
~ add Curtis fig2 params to killer figure
~ tricky adding to legend or labelling
~ can I get legend for killer figure beneath ?
~ can I derive a line for the border between when seq & mix better by fitting a line to the 0 values ?
~ get the classification tree that matches my killer figure

** 9/3/2016 **
~ creating 20% tree & plot
~ got the 0 line onto the first facet of the 20% plot

** 10/3/2016 **
~ looking at the manuscript

I think the manuscript can be about describing development and results of a model that is a more generic version of Curtis. But we do not need to make the model useable by others at this stage.

** 11/3/2016 **
manuscript

current wordcount 9,300

Details from Plos Computational Biol :
Software Submissions
PLOS Computational Biology publishes articles describing outstanding open source software of exceptional importance that has been shown to provide new biological insights, either as a part of the software article, or published elsewhere.

The software must already be widely adopted, or have the promise of wide adoption by a broad community of users. Enhancements to existing published open source software will only be considered if those enhancements bring exceptional new capabilities. The software must be downloadable anonymously in source code form and licensed under an Open Source Initiative (OSI) compliant license. The source code must be accompanied with documentation on building and installing the software from source, as well as for using the software, including instructions on how a user can test the software on supplied test data. Software articles require a presubmission inquiry that includes explanations on how the above criteria are met.

Format
Articles should be concise (less than 3500 words, not including supplementary material)

Sentence on how the decisoin tree level decided.

A rule of thumb is to choose the lowest level where the rel_error + xstd < xerror
xerror is the cross-validation error

To avoid overfitting the number of levels was restricted to the minimum where the relative error plus the standard error was less than the cross-validation error.

added mix both to paper1_results_figs.Rmd
repeated all figs for 50% resistance in paper1_results_figs_slimmed_50.Rmd

@@

In fig x1 sequential & adaptive mixture look very similar indeed ???

Could I write a very short paper describing the shiny app that allows users to take the curtis figure & modify the code ? Where might it go ? It wouldn't need any of the sensitivity analysis etc. ? It could cite the paper with the snesitivity analysis.

Repeat my killer fig for mixture both (because of slight uncertainty in the adaptive methodology)?

Tables 3 & 4 in the manuscript- can we do these as a figure ?
These address Curtis statement

Do we want to go with a 50% threshold across all analyses to keep clearer ?

An influential assertion is that mixtures may be counter-productive (made in the abstract and  repeated on page 261) where Curtis refers to his Example (vi) on his Table 1. Unfortunately, Curtis Figure 1 only shows the spread of resistance using a mixture and does not quantify the extent to which it is inferior to sequential use. However, the calibration used to replicate Curtis Figure 1 allows us to extend the one-generation argument presented in Curtis Table 1(vi) over the numerous generations required to simulate the spread of resistance (see later discussion of our Table 3).

@@

Potential for me to look at in a 2nd paper :
Differential exposure of males & females. (I have shown exposure is important so this could have a big effect).

@@

@@

@@

@@

@@
